---
title: "RAW-C: Supplemental Analysis"
author: "Sean Trott"
date: "May 5, 2024"
output:
  # pdf_document: 
  #    fig_caption: yes
  #    keep_md: yes
  #    keep_tex: yes
  html_document:
     keep_md: yes
     toc: yes
     toc_float: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300, fig.format = "pdf", warning = FALSE, message = FALSE)
```

```{r include=FALSE}
library(tidyverse)
library(lmtest)
library(forcats)
library(broom)
library(lme4)
library(ggridges)
library(lmerTest)
library(ggrepel)
library(tools)
library(viridis)
library(corrplot)
library(gplots)

all_colors <- viridis::viridis(10, option = "mako")
my_colors <- all_colors[c(3, 5, 7)]  # Selecting specific colors from the palette
my_colors <- all_colors[c(3, 7)] 
```

# Load data

```{r}
### setwd("/Users/seantrott/Dropbox/UCSD/Research/Ambiguity/SSD/spanish_norms/src/analysis/")
### Read in all data
df_rawc = read_csv("../../data/extra/rawc_stimuli.csv")
```



# Comparing all models together

```{r}
df_bert = read_csv("../../data/processed/models_english/rawc-distances_model-bert-base-uncased.csv") %>%
  mutate(Model = "BERT-base-uncased",
         Multilingual = "Monolingual")

df_bert_cased = read_csv("../../data/processed/models_english/rawc-distances_model-bert-base-cased.csv") %>%
  mutate(Model = "BERT-base-cased",
         Multilingual = "Monolingual")


df_xlm = read_csv("../../data/processed/models_english/rawc-distances_model-xlm-roberta-base.csv") %>%
  mutate(Model = "XLM-RoBERTa",
         Multilingual = "Multilingual")


df_ab1 = read_csv("../../data/processed/models_english/rawc-distances_model-albert-base-v1.csv") %>%
  mutate(Model = "ALBERT-base-v1",
         Multilingual = "Monolingual")

df_ab2 = read_csv("../../data/processed/models_english/rawc-distances_model-albert-base-v2.csv") %>%
  mutate(Model = "ALBERT-base-v2",
         Multilingual = "Monolingual")

df_al = read_csv("../../data/processed/models_english/rawc-distances_model-albert-large-v2.csv") %>%
  mutate(Model = "ALBERT-large-v2",
         Multilingual = "Monolingual")

df_axl = read_csv("../../data/processed/models_english/rawc-distances_model-albert-xlarge-v2.csv") %>%
  mutate(Model = "ALBERT-xlarge-v2",
         Multilingual = "Monolingual")

df_axxl = read_csv("../../data/processed/models_english/rawc-distances_model-albert-xxlarge-v2.csv")  %>%
  mutate(Model = "ALBERT-xxlarge-v2",
         Multilingual = "Monolingual")

df_rb = read_csv("../../data/processed/models_english/rawc-distances_model-roberta-base.csv") %>%
  mutate(Model = "RoBERTa-base",
         Multilingual = "Monolingual")

df_rl = read_csv("../../data/processed/models_english/rawc-distances_model-roberta-large.csv") %>%
  mutate(Model = "RoBERTa-large",
         Multilingual = "Monolingual")

df_db = read_csv("../../data/processed/models_english/rawc-distances_model-distilbert-base-uncased.csv") %>%
  mutate(Model = "DistilBERT",
         Multilingual = "Monolingual")


df_mb = read_csv("../../data/processed/models_english/rawc-distances_model-bert-base-multilingual-cased.csv") %>%
  mutate(Model = "Multilingual BERT",
         Multilingual = "Multilingual")



df_all = df_bert %>%
  bind_rows(df_bert_cased) %>%
  bind_rows(df_xlm) %>%
  bind_rows(df_ab1) %>%
  bind_rows(df_ab2) %>%
  bind_rows(df_al) %>%
  bind_rows(df_axl) %>%
  bind_rows(df_axxl) %>%
  bind_rows(df_rb) %>%
  bind_rows(df_rl) %>%
  bind_rows(df_db) %>%
  bind_rows(df_mb)

nrow(df_all)


### merge together
df_merged = df_all %>%
  inner_join(df_rawc, by = c("sentence1", "sentence2", "word", "string"))

```


## Same/Different by layer, by model

```{r}
# Calculate mean and SE of Distance
df_summary <- df_merged %>%
  group_by(Layer, Same_sense, Model) %>%
  summarise(
    mean_Distance = mean(Distance),
    sd_Distance = sd(Distance),
    count = n(),
    se_Distance = sd_Distance / sqrt(count),
    .groups = 'drop'  # Drop the automatic grouping by dplyr
  )

df_summary %>%
  ggplot(aes(x = Layer, 
             y = mean_Distance, 
             color = Same_sense, 
             fill = Same_sense)) +
  geom_line(size = 1.5) +
  geom_ribbon(aes(ymin = mean_Distance - 2 * se_Distance, 
                  ymax = mean_Distance + 2 * se_Distance), 
              alpha = 0.8,
              color = NA) +
  labs(x = "Layer",
       y = "Mean Distance",
       color = "Same Sense",
       fill = "Same Sense") +
  # scale_x_continuous(breaks = c(0:13)) +
  theme_minimal() +
  # scale_fill_viridis(option = "mako", discrete=TRUE) +
  # scale_color_viridis(option = "mako", discrete=TRUE) +
  scale_fill_manual(values = my_colors)  +
  scale_color_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="bottom") +
  facet_wrap(~Model)

```


## R2 by layer, by model

```{r r2_by_layer}
df_by_layer = df_merged %>%
  group_by(Model, Multilingual, Layer, n_params) %>%
  summarise(r = cor(mean_relatedness, Distance, method = "pearson"),
            r2 = r ** 2,
            rho = cor(mean_relatedness, Distance, method = "spearman"),
            count = n())

max(df_by_layer$r2)
df_by_layer %>%
  filter(r2 == max(df_by_layer$r2))

df_by_layer %>%
  filter(Model != "ALBERT-tiny") %>%
  filter(Model != "ALBERT-xlarge-v2") %>%
  filter(Model != "DistilBETO") %>%
  filter(Model != "ALBERT-large-v2") %>%
  filter(Model != "RoBERTa-large") %>%
  ggplot(aes(x = Layer,
             y = r2,
             color = Model,
             linetype = Model)) +
  geom_line(size = 2) +
  theme_minimal() +
  labs(x = "Layer",
       y = "R2",
       color = "",
       linetype = "") +
  scale_x_continuous(breaks = c(0:13)) +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 12),
        legend.position="bottom")

```

We can also break this down by looking at the *proportion of layers* through the model.

```{r r2_by_prop_layer}
df_families = data.frame(
  Model = unique(df_by_layer$Model),
  Family = c("ALBERT", "ALBERT", "ALBERT", "ALBERT","ALBERT",
             "BERT", "BERT","BERT", "BERT",
             "RoBERTa", "RoBERTa", "RoBERTa"))

df_by_layer = df_by_layer %>%
  group_by(Model) %>%
  mutate(max_layer = max(Layer),
         prop_layer = Layer / max_layer)  %>%
  inner_join(df_families, by = "Model")


df_by_layer %>%
  mutate(binned_prop_layer = ntile(prop_layer, 10)) %>%
  mutate(prop_binned = binned_prop_layer / 10) %>%
  ggplot(aes(x = prop_binned,
             y = r2)) +
  geom_line(size = .5, aes(color = Model, group = Model),
            alpha = .3) + 
  stat_summary(
    fun = mean,                    # calculate mean
    geom = "line",                 # use line geometry
    size = 2,                    # set line size
    na.rm = TRUE                   # remove NA values
  ) +
  theme_minimal() +
  labs(x = "Layer Depth Ratio",
       y = "R2",
       color = "") +  # Renaming legend title appropriately
  scale_color_viridis(option = "mako", discrete = TRUE) +
  theme(text = element_text(size = 15),
        legend.position = "none") 

  
df_by_layer %>%
  mutate(binned_prop_layer = ntile(prop_layer, 10)) %>%
  mutate(prop_binned = binned_prop_layer / 10) %>%
  ggplot(aes(x = prop_binned,
             y = r2)) +
  stat_summary(
    aes(group = Family,
        color = Family),  
    fun = mean,    
    geom = "line",        
    size = 2              
  ) +
  stat_summary(
    aes(group = Family, 
        fill = Family), 
    fun.data = mean_se,    
    geom = "ribbon",  
    alpha = 0.2,   
    color = NA     
  ) +
  theme_minimal() +
  labs(x = "Layer Depth Ratio",
       y = "R2",
       color = "Family") +
  scale_color_viridis(option = "mako", discrete = TRUE) +
  theme(text = element_text(size = 15),
        legend.position = "bottom") 
```


## Comparison by model parameters

The BETO/ALBERT/DistilBETO family of models provide a nice within-family comparison of the effect of number of **parameters**, since they are trained on the same datasets.

```{r r2_by_params}

df_max_r2 = df_by_layer %>%
  group_by(Model, n_params, Multilingual) %>%
  summarise(max_r2 = max(r2))

df_max_r2 %>%
  ggplot(aes(x = n_params,
             y = max_r2,
             color = Multilingual,
             shape = Multilingual)) +
  geom_point(size = 6,
             alpha = .7) +
  geom_hline(yintercept = .64,
              linetype = "dotted", color = "red",
             size = 1.2, alpha = .5) +
  geom_text_repel(aes(label=Model), size=3) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Parameters",
       y = "Maximum R2",
       color = "",
       title = "Scaling on English RAW-C Dataset",
       shape = "") +
  theme_minimal() +
  # guides(color="none") +
  # scale_color_viridis(option = "mako", discrete=TRUE) +
  scale_color_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="bottom")


summary(lm(data = df_max_r2,
           max_r2 ~ log10(n_params) + Multilingual))

```

## `Same` vs. `Different` Sense residuals


```{r same_diff_residuals}
# Fit models and calculate residuals for each LLM
models <- by(df_merged, df_merged$Model, function(subdata) {
  model <- lm(mean_relatedness ~ Distance * Layer, data = subdata)
  subdata$residuals <- residuals(model)
  return(subdata)
})

# Combine the results back into a single dataframe
results <- do.call(rbind, models)

# Modify Same_sense
results = results %>%
  mutate(Same_sense = case_when(
    Same_sense == TRUE ~ "Same Sense",
    Same_sense == FALSE ~ "Different Sense"
  ))
  



results %>%
  ggplot(aes(x = residuals,
             fill = Same_sense,
             y = ambiguity_type)) +
  # geom_density(alpha = .7, size = 0) +
  geom_density_ridges2(aes(height = ..density..), 
                       color=gray(0.25), 
                       alpha = .7, 
                       scale=.85, 
                       # size=1, 
                       size = 0,
                       stat="density") +
  geom_vline(xintercept = 0, size = .6, linetype = "dashed") +
  theme_minimal() +
  # scale_fill_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position="bottom") +
  scale_fill_manual(values = my_colors)  +
  labs(x = "Residuals (Relatedness ~ Distance * Layer)",
       fill = "",
       y = "Ambiguity Type") +
  facet_wrap(~Model)

```

