---
title: "SAW-C: Main Relatedness Analysis"
author: "Sean Trott"
date: "April 11, 2024"
output:
  # pdf_document: 
  #    fig_caption: yes
  #    keep_md: yes
  #    keep_tex: yes
  html_document:
     keep_md: yes
     toc: yes
     toc_float: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300, fig.format = "pdf", warning = FALSE, message = FALSE)
```

```{r include=FALSE}
library(tidyverse)
library(lmtest)
library(forcats)
library(broom)
library(lme4)
library(ggridges)
library(lmerTest)
library(ggrepel)
library(tools)
library(viridis)

all_colors <- viridis::viridis(10, option = "mako")
my_colors <- all_colors[c(3, 5, 7)]  # Selecting specific colors from the palette
my_colors <- all_colors[c(3, 7)] 
```

# Load data

```{r}
### setwd("/Users/seantrott/Dropbox/UCSD/Research/Ambiguity/SSD/spanish_norms/src/analysis/")
### Read in all data
df_final = read_csv("../../data/processed/human/sawc_relatedness_full_critical_data.csv")
nrow(df_final)
length(unique(df_final$Participant))
```


# RQ1: Same vs. Different Sense

```{r rq1_same_different}
df_ratio = df_final %>% 
  group_by(Same_sense) %>%
  mutate(count_condition = n()) %>%
  ungroup() %>%
  group_by(Same_sense, Response, count_condition) %>%
  summarise(count_response = n()) %>%
  mutate(prop_response = count_response / count_condition)
              
df_ratio %>%         
  ggplot(aes(x = Response,
             y = prop_response)) +
  geom_bar(alpha = .6, stat = "identity") +
  theme_minimal() +
  labs(x = "Relatedness",
       y = "P(Response | Condition)") +
  facet_wrap(~Same_sense) +
  theme(text = element_text(size = 15),
        legend.position="none")

df_final %>%         
  ggplot(aes(x = Response)) +
  geom_bar(alpha = .6, stat = "count") +
  theme_minimal() +
  labs(x = "Relatedness",
       y = "Count") +
  facet_wrap(~Same_sense) +
  theme(text = element_text(size = 15),
        legend.position="none")


mod_full = lmer(data = df_final,
                Response ~ Same_sense +
                  (1 + Same_sense | Participant) + 
                  (1 + Same_sense | List) + (1 | Word),
                REML = FALSE)

mod_reduced = lmer(data = df_final,
                Response ~ # Same_sense +
                  (1 + Same_sense | Participant) + 
                  (1 + Same_sense | List) + (1 | Word),
                REML = FALSE)

summary(mod_full)
anova(mod_full, mod_reduced)
```


# RQ2 + RQ5: Correlation with BETO

## Load and process BETO data

```{r}
### BETO distances
df_beto_distances = read_csv("../../data/processed/models/sawc-distances_model-bert-base-spanish-wwm-cased.csv")
nrow(df_beto_distances)

### basic items to get Tag
df_sawc_items = read_csv("../../data/raw/items/sawc_sentence_pairs.csv")

### merge together
df_merged = df_beto_distances %>%
  left_join(df_sawc_items)
nrow(df_merged)
```

## Merge with SAW-C Norms

```{r}
df_list_mean = df_final %>%
  group_by(List, Word, Tag) %>%
  summarise(mean_relatedness = mean(Response), .groups = "drop",
            count = n())
nrow(df_list_mean)

df_merged_beto = df_merged %>%
  inner_join(df_list_mean)
nrow(df_merged_beto)
```

## RQ2: Correlation by layer

```{r rq2_corr_by_layer}
df_by_layer = df_merged_beto %>%
  group_by(Layer) %>%
  summarise(r = cor(mean_relatedness, Distance, method = "pearson"),
            r2 = r ** 2,
            rho = cor(mean_relatedness, Distance, method = "spearman"),
            count = n())

summary(df_by_layer$rho)

df_by_layer %>%
  ggplot(aes(x = Layer,
             y = r)) +
  geom_line(size = 2,
            color = "cornflower blue") +
  theme_minimal() +
  labs(x = "Layer (BETO)",
       y = "Pearson's r") +
  scale_x_continuous(breaks = c(0:13)) +
  theme(text = element_text(size = 15),
        legend.position="none")


df_by_layer %>%
  ggplot(aes(x = Layer,
             y = rho)) +
  geom_line(size = 2,
            color = "cornflower blue") +
  theme_minimal() +
  labs(x = "Layer (BETO)",
       y = "Spearman's rho") +
  scale_x_continuous(breaks = c(0:13)) +
  theme(text = element_text(size = 15),
        legend.position="none")

df_by_layer %>%
  ggplot(aes(x = Layer,
             y = r2)) +
  geom_line(size = 2,
            color = "cornflower blue") +
  theme_minimal() +
  labs(x = "Layer (BETO)",
       y = "R2") +
  scale_x_continuous(breaks = c(0:13)) +
  theme(text = element_text(size = 15),
        legend.position="none")

min(df_by_layer$rho)
```


## RQ5: Expected layer

```{r rq5_layer}
df_by_layer <- df_by_layer %>%
  mutate(r2_delta = c(NA, diff(r2))) %>%
  mutate(weighted_layer = Layer * r2_delta)

expected_layer = sum(df_by_layer$weighted_layer, na.rm = TRUE) / sum(df_by_layer$r2_delta, na.rm = TRUE)
expected_layer

df_by_layer %>%
  ggplot(aes(x = Layer,
             y = r2)) +
  geom_line(size = 2,
            color = "cornflower blue",
            alpha = .8) +
  theme_minimal() +
  geom_vline(xintercept = expected_layer, size = 1.5, 
             linetype = "dashed", alpha = .7) +
  geom_vline(xintercept = 7, size = 1.5, 
             linetype = "dotted", alpha = .7) +
  scale_x_continuous(breaks = c(0:13)) +
  labs(x = "Layer (BETO)",
       y = "R2") +
  theme(text = element_text(size = 15),
        legend.position="none") 
```




# RQ3: Cosine distance vs. Same/Different

Now, we select the *best-performing layer* from BETO.

```{r rq3}
df_beto_l5 = df_merged %>%
  filter(Layer == 7) %>%
  select(-Same_sense)
nrow(df_beto_l5)

df_experimental_with_beto = df_final %>%
  left_join(df_beto_l5)
nrow(df_experimental_with_beto)
  

mod_full = lmer(data = df_experimental_with_beto,
                Response ~ Same_sense + Distance +
                  (1 + Same_sense + Distance | Participant) + 
                  (1 | List) + (1 | Word),
                REML = FALSE)

mod_reduced = lmer(data = df_experimental_with_beto,
                Response ~ Distance + # Same_sense +
                  (1 + Same_sense + Distance | Participant) + 
                  (1 | List) + (1 | Word),
                REML = FALSE)

mod_just_same = lmer(data = df_experimental_with_beto,
                Response ~ Same_sense + # Distance
                  (1 + Same_sense + Distance | Participant) + 
                  (1 | List) + (1 | Word),
                REML = FALSE)

summary(mod_full)
anova(mod_full, mod_reduced)
anova(mod_full, mod_just_same)



### Visualize
df_experimental_with_beto %>%
  mutate(distance_binned = ntile(Distance, 20)) %>%
  group_by(Same_sense, distance_binned) %>%
  summarize(
    mean_relatedness = mean(Response),
    sd_relatedness = sd(Response),
    count = n(),
    se_relatedness = sd_relatedness / sqrt(count),
  ) %>%
  ggplot(aes(x = distance_binned, 
             y = mean_relatedness, 
             color = Same_sense, 
             fill = Same_sense)) +
  geom_line(size = 1.5) +
  geom_ribbon(aes(ymin = mean_relatedness - se_relatedness, 
                  ymax = mean_relatedness + se_relatedness), 
              alpha = 0.8,
              color = NA) +
  labs(x = "BETO Cosine Distance (Binned)",
       y = "Relatedness",
       color = "Same Sense",
       fill = "Same Sense") +
  theme_minimal() +
  scale_fill_manual(values = my_colors)  +
  scale_color_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="bottom")
  
```



# RQ4: BETO vs. inter-annotator agreement

```{r rq4_comparison}
### First, get group means
df_list_mean = df_experimental_with_beto %>%
  group_by(List, Word, Tag) %>%
  summarise(mean_relatedness = mean(Response), .groups = "drop",
            count = n(),
            distance = mean(Distance))
nrow(df_list_mean)

### Get BETO cor
BETO_COR = abs(cor(df_list_mean$mean_relatedness,
               df_list_mean$distance, method = "spearman"))

### Now, iterate through ppts
ppts = unique(df_final$Participant)
df_r = data.frame()

for (ppt in ppts) {
  # Subset df_critical for the current participant
  individual_data <- df_final %>%
    filter(Participant == ppt) %>%
    select(List, Word, Tag, Response)  # Ensure you're selecting the needed columns

  # Merge individual data with mean data
  merged_data <- inner_join(individual_data, df_list_mean, by = c("List", "Word", "Tag"))
  
  test = cor.test(merged_data$Response,
                  merged_data$mean_relatedness,
                 method = "spearman")
  
  df_test = broom::tidy(test)
  df_test$ppt = ppt
  df_test$List = unique(individual_data$List)
  df_r = rbind(df_r, df_test)

}

### Visualization
df_r %>%
  ggplot(aes(x = estimate)) +
  geom_histogram(alpha = .6) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_minimal() +
  geom_vline(xintercept = BETO_COR, size = 1.5, 
             linetype = "dashed", alpha = .7) +
  labs(x = "Leave-one-out Correlation") +
  theme(text = element_text(size = 15),
        legend.position="none")

summary(df_r$estimate)


### What proportion of inter-annotator agreement scores are larger?
prop_larger = df_r %>%
  mutate(larger = estimate >= BETO_COR) %>%
  summarise(mean(larger))
prop_larger


```


# RQ6: Layer vs. same/different

The actual code implementation is executed in Python, but a visualization is produced here.

```{r rq6_same_different}
# Calculate mean and SE of Distance
df_summary <- df_beto_distances %>%
  group_by(Layer, Same_sense) %>%
  summarise(
    mean_Distance = mean(Distance),
    sd_Distance = sd(Distance),
    count = n(),
    se_Distance = sd_Distance / sqrt(count),
    .groups = 'drop'  # Drop the automatic grouping by dplyr
  )

df_summary %>%
  ggplot(aes(x = Layer, 
             y = mean_Distance, 
             color = Same_sense, 
             fill = Same_sense)) +
  geom_line(size = 1.5) +
  geom_ribbon(aes(ymin = mean_Distance - 2 * se_Distance, 
                  ymax = mean_Distance + 2 * se_Distance), 
              alpha = 0.8,
              color = NA) +
  labs(x = "Layer (BETO)",
       y = "Mean Distance",
       color = "Same Sense",
       fill = "Same Sense") +
  scale_x_continuous(breaks = c(0:13)) +
  theme_minimal() +
  # scale_fill_viridis(option = "mako", discrete=TRUE) +
  # scale_color_viridis(option = "mako", discrete=TRUE) +
  scale_fill_manual(values = my_colors)  +
  scale_color_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="bottom")
  
```


# Additional visualizations

```{r visualizations}
df_item_means = df_final %>%
  group_by(List, Word, Same_sense, Sentence_1, Sentence_2,
           Sense_id_s1, Sense_id_s2, Gender_s1, Gender_s2) %>%
  summarise(mean_relatedness = mean(Response),
            sd_relatedness = sd(Response),
            median_relatedness = median(Response),
            count = n())


df_item_means %>%         
  ggplot(aes(x = mean_relatedness)) +
  geom_histogram(alpha = .6, bins = 10) +
  theme_minimal() +
  labs(x = "Mean Relatedness",
       y = "Count") +
  facet_wrap(~Same_sense) +
  theme(text = element_text(size = 15),
        legend.position="none")

df_item_means %>%
  ggplot(aes(x = mean_relatedness,
             y = Same_sense,
             fill = Same_sense)) +
  geom_density_ridges2(aes(height = ..density..), 
                       color=gray(0.25), 
                       alpha = .7, 
                       scale=.85, 
                       # size=1, 
                       size = 0,
                       stat="density") +
  labs(x = "Mean Relatedness",
       y = "",
       fill = "") +
  theme_minimal()+
  scale_fill_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="none") 

```




# Comparing all models together

```{r}
df_beto = read_csv("../../data/processed/models/sawc-distances_model-bert-base-spanish-wwm-cased.csv") %>%
  mutate(Model = "BETO-cased",
         Multilingual = "Monolingual")
df_xlm = read_csv("../../data/processed/models/sawc-distances_model-xlm-roberta-base.csv") %>%
  mutate(Model = "XLM-RoBERTa",
         Multilingual = "Multilingual")

df_mb = read_csv("../../data/processed/models/sawc-distances_model-bert-base-multilingual-cased.csv") %>%
  mutate(Model = "Multilingual BERT",
         Multilingual = "Multilingual")

df_db = read_csv("../../data/processed/models/sawc-distances_model-distilbert-base-spanish-uncased.csv") %>%
  mutate(Model = "DistilBETO",
         Multilingual = "Monolingual")

df_ab = read_csv("../../data/processed/models/sawc-distances_model-albert-base-spanish.csv") %>%
  mutate(Model = "ALBERT-base",
         Multilingual = "Monolingual")

df_at = read_csv("../../data/processed/models/sawc-distances_model-albert-tiny-spanish.csv") %>%
  mutate(Model = "ALBERT-tiny",
         Multilingual = "Monolingual")

df_axl = read_csv("../../data/processed/models/sawc-distances_model-albert-xlarge-spanish.csv") %>%
  mutate(Model = "ALBERT-xlarge",
         Multilingual = "Monolingual")

df_al = read_csv("../../data/processed/models/sawc-distances_model-albert-large-spanish.csv") %>%
  mutate(Model = "ALBERT-large",
         Multilingual = "Monolingual")

df_axxl = read_csv("../../data/processed/models/sawc-distances_model-albert-xxlarge-spanish.csv") %>%
  mutate(Model = "ALBERT-xxlarge",
         Multilingual = "Monolingual")

df_rb = read_csv("../../data/processed/models/sawc-distances_model-roberta-base-bne.csv") %>%
  mutate(Model = "RoBERTa-base",
         Multilingual = "Monolingual")

df_rl = read_csv("../../data/processed/models/sawc-distances_model-roberta-large-bne.csv") %>%
  mutate(Model = "RoBERTa-large",
         Multilingual = "Monolingual")

df_beto_uncased = read_csv("../../data/processed/models/sawc-distances_model-bert-base-spanish-wwm-uncased.csv") %>%
  mutate(Model = "BETO-uncased",
         Multilingual = "Monolingual")

df_all = df_beto %>%
  bind_rows(df_xlm) %>%
  bind_rows(df_db) %>%
  bind_rows(df_ab) %>%
  bind_rows(df_at) %>%
  bind_rows(df_mb) %>%
  bind_rows(df_axl) %>%
  bind_rows(df_al) %>%
  bind_rows(df_axxl) %>%
  bind_rows(df_rb) %>%
  bind_rows(df_rl) %>%
  bind_rows(df_beto_uncased)

nrow(df_all)


### basic items to get Tag
df_sawc_items = read_csv("../../data/raw/items/sawc_sentence_pairs.csv")

### merge together
df_merged = df_all %>%
  left_join(df_sawc_items)

df_list_mean = df_final %>%
  group_by(List, Word, Tag) %>%
  summarise(mean_relatedness = mean(Response), .groups = "drop",
            count = n())
nrow(df_list_mean)

df_merged = df_merged %>%
  inner_join(df_list_mean)
nrow(df_merged)
```


## R2 by layer, by model

```{r r2_by_layer}
df_by_layer = df_merged %>%
  group_by(Model, Multilingual, Layer, n_params) %>%
  summarise(r = cor(mean_relatedness, Distance, method = "pearson"),
            r2 = r ** 2,
            rho = cor(mean_relatedness, Distance, method = "spearman"),
            count = n())

max(df_by_layer$r2)
df_by_layer %>%
  filter(r2 == max(df_by_layer$r2))

df_by_layer %>%
  filter(Model != "ALBERT-tiny") %>%
  filter(Model != "ALBERT-xlarge") %>%
  filter(Model != "DistilBETO") %>%
  filter(Model != "ALBERT-large") %>%
  filter(Model != "RoBERTa-large") %>%
  ggplot(aes(x = Layer,
             y = r2,
             color = Model,
             linetype = Model)) +
  geom_line(size = 2) +
  theme_minimal() +
  labs(x = "Layer",
       y = "R2",
       color = "",
       linetype = "") +
  scale_x_continuous(breaks = c(0:13)) +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 12),
        legend.position="bottom")

df_by_layer %>%
  filter(Model != "ALBERT-tiny") %>%
  filter(Model != "ALBERT-xlarge") %>%
  filter(Model != "DistilBETO") %>%
  filter(Model != "ALBERT-large") %>%
  filter(Model != "RoBERTa-large") %>%
  ggplot(aes(x = Layer,
             y = r2,
             color = Model)) +
  geom_line(size = 2) +
  theme_minimal() +
  labs(x = "Layer",
       y = "R2",
       color = "") +
  scale_x_continuous(breaks = c(0:13)) +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 12),
        legend.position="bottom")

df_by_layer %>%
  filter(Model %in% c("ALBERT-xlarge", 
                      "ALBERT-large",
                      "RoBERTa-large")) %>%
  ggplot(aes(x = Layer,
             y = r2,
             color = Model)) +
  geom_line(size = 2) +
  theme_minimal() +
  labs(x = "Layer",
       y = "R2",
       color = "") +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position="bottom")


```

## Comparison by model parameters

The BETO/ALBERT/DistilBETO family of models provide a nice within-family comparison of the effect of number of **parameters**, since they are trained on the same datasets.

```{r r2_by_params}

df_max_r2 = df_by_layer %>%
  group_by(Model, n_params, Multilingual) %>%
  summarise(max_r2 = max(r2))

df_max_r2 %>%
  ggplot(aes(x = n_params,
             y = max_r2,
             color = Multilingual,
             shape = Multilingual)) +
  geom_point(size = 6,
             alpha = .7) +
  geom_hline(yintercept = mean(df_r$estimate)**2, ### Human accuracy 
              linetype = "dotted", color = "red",
             size = 1.2, alpha = .5) +
  geom_text_repel(aes(label=Model), size=3) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Parameters",
       y = "Maximum R2",
       color = "",
       shape = "") +
  theme_minimal() +
  # guides(color="none") +
  # scale_color_viridis(option = "mako", discrete=TRUE) +
  scale_color_manual(values = my_colors)  +
  theme(text = element_text(size = 15),
        legend.position="bottom")

```

## `Same` vs. `Different` Sense residuals


```{r same_diff_residuals}
# Fit models and calculate residuals for each LLM
models <- by(df_merged, df_merged$Model, function(subdata) {
  model <- lm(mean_relatedness ~ Distance * Layer, data = subdata)
  subdata$residuals <- residuals(model)
  return(subdata)
})

# Combine the results back into a single dataframe
results <- do.call(rbind, models)

# Modify Same_sense
results = results %>%
  mutate(Same_sense = case_when(
    Same_sense == TRUE ~ "Same Sense",
    Same_sense == FALSE ~ "Different Sense"
  ))
  



results %>%
  ggplot(aes(x = residuals,
             fill = Same_sense)) +
  geom_density(alpha = .7, size = 0) +
  geom_vline(xintercept = 0, size = .6, linetype = "dashed") +
  theme_minimal() +
  # scale_fill_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position="bottom") +
  scale_fill_manual(values = my_colors)  +
  labs(x = "Residuals (Relatedness ~ Distance * Layer)",
       fill = "") +
  facet_wrap(~Model)

```

